---
title: "Fundamentals - 2"
author: "Abhishek Dixit"
date: "3/3/2023"
output:
  pdf_document:
    latex_engine: xelatex
header-includes:
  \usepackage{fvextra}
  \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

# Question 1: (50 points)

## 1.1 
Find the attached JPM.csv file. Use as.Date() function to change the first column to Date object.

```{r}
setwd('C:\\Users\\amand\\OneDrive\\Desktop\\FE 515 (R)\\HW_2')
df <- read.csv(file='JPM.csv', header=TRUE, sep=",", na.string = "?")
df[,1] <- as.Date(df[,1], origin="2010-01-01")
#view(df)

```

## 1.2

Plot the adjusted close price against the date object (i.e. date object as
x-axis and close price as y-axis) in red line (require no points). Set the
title as JPM, the label for x-axis as Date and the label for y-axis as
Adjusted Close Price.

```{r}
plot(df$X, df$JPM.Adjusted, type = "l", main = "JPM", xlab = "Date", ylab = "Adjusted Close Price", col = "red")
```

## 1.3 

Create a scatter plot of close price against open price (i.e. open prices as
x-axis, and close prices as y-axis). Set the x label as ”Open Price” and y
label as ”Close Price”.

```{r}
plot(df$JPM.Open, df$JPM.Close, main = "Scatter Plot", xlab = "Open Price", ylab = "Close Price")
```

## 1.4 

Use cut() function to divide adjusted close price into 4 intervals. Generate
a barplot for the frequencies of these intervals.


```{r}
barplot(table(cut(df$JPM.Adjusted, breaks = 4)), xlab = "Adjusted Close Price", ylab = "Frequency")
```

## 1.5

Generate a boxplot of volume against the 4 intervals of adjusted close
prices.

```{r}
boxplot(df$JPM.Volume ~ cut(df$JPM.Adjusted, breaks = 4), xlab = "Adjusted Close Price", ylab = "Volume")
```

## 1.6

Use par() function to create a picture of 4 subplots. Gather the 4 figures
from 1.2 - 1.5 into ONE single picture. Please arrange the 4 subplots
into a 2 by 2 frame, i.e. a frame consists of 2 columns and 2 rows. (Hint.
par(mfrow = c(1,3)) will create a picture of three subplots. In the picture, the subplots are arranged into a 1 by 3 frame.)


```{r}
par(mfrow = c(2,2))
plot(df$X, df$JPM.Adjusted, type = "l", main = "JPM", xlab = "Date", ylab = "Adjusted Close Price", col = "red")
plot(df$JPM.Open, df$JPM.Close, main = "Scatter Plot", xlab = "Open Price", ylab = "Close Price")
barplot(table(cut(df$JPM.Adjusted, breaks = 4)), xlab = "Adjusted Close Price", ylab = "Frequency")
boxplot(df$JPM.Volume ~ cut(df$JPM.Adjusted, breaks = 4), xlab = "Adjusted Close Price", ylab = "Volume")

```

# Question 2

Estimate the volume of the unit sphere (which is just 4π/3) by simulation.

```{r}
seed <- 42 # let the seed be 42 in global environment

rnd <- function(n){
m <- 2 ^ 31 - 1
a <- 7 ^ 5
b <- 0
x <- rep(NA, n)
x[1] <- (a * seed + b) %% m
for(i in 1:(n-1)){
x[i + 1] <- (a * x[i] + b) %% m
}
seed <<- x[n] # change the seed in global environment
return(x/m)
}

N <- 10000
x <- rnd(N)
y <- rnd(N)
z <- rnd(N)
n.red.bool <- x^2 + y^2 + z^2 <= 1
n.red.num <- sum(n.red.bool) # vectorized calculation

volume.pie.sphere <- n.red.num / N
total_volume <- 8 * volume.pie.sphere
total_volume
```

# Question 3

## 3.1 

Implement a Linear Congruential Generator (LCG) which generates pseudo-random number from uniform distribution using m = 244944, a = 1597, b = 51749.


```{r}
lcg <- function(n=5){
  v <- vector(length = n)
  m <- 244944
  a <- 1597
  b <- 51749
  d <- as.numeric(Sys.time()) * 1000
  for (i in 1:n){
    d <- (a * d + b) %% m
    v[i] <- d / m
    }
  return (v)
}
lcg(5)
```

## 3.2

Use the LCG in the previous problem, generate 10000 random numbers from chi-square distribution with 10 degrees of freedom (i.e. df = 10), and assign to a variable.

```{r}
chi_sq <- qchisq(lcg(n=10000), 10)
#chi_sq
```

## 3.3

Visualize the resulting sample from 3.2 using a histogram with 40 bins.

```{r}
hist(chi_sq, breaks=40)
```
