---
title: "2023S FE 515-B Assignment 3"
author: "Abhishek Dixit"
date: "4/21/2023"
output:
  pdf_document:
    latex_engine: xelatex
header-includes:
  \usepackage{fvextra}
  \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

# Question 1: (50 points)

## 1.1 
Download option prices of ticker VIX for all expiration dates and name
it VIX.options

```{r}
library(quantmod)

VIX.options <- getOptionChain('^VIX',NULL)

head(VIX.options)

```

## 1.2
Download the current price (last quote price) for ˆVIX

```{r}
current_price <- getQuote('^VIX')$Last
current_price
```

## 1.3 
For calls and puts of VIX.options at each expiration calculate the average
of Bid and Ask. Create a new column named ‘Price‘ to contain the result.

```{r}
for (i in 1:length(VIX.options)){
	VIX.options[[i]]$calls$Price <- (VIX.options[[i]]$calls$Bid + VIX.options[[i]]$calls$Ask)/2
	VIX.options[[i]]$puts$Price <- (VIX.options[[i]]$puts$Bid + VIX.options[[i]]$puts$Ask)/2
	}
	
head(VIX.options)
```

## 1.4 
For calls and puts of VIX.options at each expiration, add a column of InTheMoney, which takes value TRUE when it is in-the-money, and FALSE otherwise. Compare it to ITM column to check your results.

```{r}
for (i in 1:length(VIX.options)){
	VIX.options[[i]]$calls$In_The_Money <- VIX.options[[i]]$calls$Strike < current_price
	VIX.options[[i]]$puts$In_The_Money <- VIX.options[[i]]$puts$Strike > current_price
}

# Compare InTheMoney and ITM columns
identical(VIX.options$InTheMoney, VIX.options$ITM)

```

## 1.5
For calls and puts of VIX at each expiration, delete all the fields except
Strike, Bid, Ask, Price, and In-The-Money, and save them in .csv files with
the format ”VIXdata2021-09-26Exp2021-10-08puts.csv”.

```{r}
for (i in 1:length(VIX.options)){
	VIX.options[[i]]$calls <- VIX.options[[i]]$calls[c('Strike', 'Bid', 'Ask', 'Price', 'In_The_Money' )]
	VIX.options[[i]]$puts <- VIX.options[[i]]$puts[c('Strike', 'Bid', 'Ask', 'Price', 'In_The_Money' )]
}

Exp <- names(VIX.options)
today <- format(Sys.Date())

setwd('C:\\Users\\amand\\OneDrive\\Desktop\\FE 515 (R)\\HW_3')

for(i in 1:length(Exp)){
	write.csv(VIX.options[[i]]$calls,file=paste("VIXdata",today,"Exp",Exp[i],"calls.csv",sep=""))
	write.csv(VIX.options[[i]]$puts,file=paste("VIXdata",today,"Exp",Exp[i],"puts.csv",sep=""))
	} 
```


# Question 2: (50 points)

## 2.1 
Using Monte-Carlo Simulation to estimate the put option price using S0 = 100, K = 100, T = 1, σ = 0.2, r = 0.05, you can use number of steps n = 252 and number of paths m = 10000

```{r}
S0 <- 100
K <- 100
T <- 1
sigma <- 0.2
r <- 0.05
n <- 252
m <- 10000

set.seed(123)
z <- matrix(rnorm(n*m, mean = 0, sd = 1), ncol = n)

dt <- T/n
expdt <- exp(-r*dt)
S <- matrix(0, nrow = m, ncol = n+1)
S[, 1] <- S0
for (i in 1:m) {
  for (j in 2:(n+1)) {
    S[i, j] <- S[i, j-1] * exp((r-0.5*sigma^2)*dt + sigma*sqrt(dt)*z[i, j-1])
  }
}

p <- matrix(0, nrow = m, ncol = 1)
for (i in 1:m) {
  p[i] <- exp(-r*T) * max(K - S[i, n+1], 0)
}
mc_price <- mean(p)

mc_price
```


## 2.2 
Implement Black-Scholes formula for pricing the put option

Check the difference between the Black-Scholes price and the Monte-Carlo price.

```{r}
d1 <- (log(S0/K) + (r + 0.5*sigma^2)*T)/(sigma*sqrt(T))
d2 <- d1 - sigma*sqrt(T)
bs_price <- -S0*pnorm(-d1) + exp(-r*T)*K*pnorm(-d2)

difference <- abs(bs_price - mc_price)
difference
```
