---
title: "2023S FE 515-B Final"
author: "Abhishek Dixit"
date: "5/1/2023"
output:
  pdf_document:
    latex_engine: xelatex
header-includes:
  \usepackage{fvextra}
  \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

# Question 1: (30 points)

## 1.1

Download the historical prices for the ticker "SPY" from 2019-01-01 until now.

```{r}
library(quantmod)

getSymbols("SPY", from = "2019-01-01")
head(SPY)
```

## 1.2

Calculate the daily log returns for SPY using the adjusted close prices.

```{r}
SPY.price <- SPY$SPY.Adjusted
SPY.log.price <- log(SPY.price)
SPY.log.return <- diff(SPY.log.price)
head(SPY.log.return)
```

## 1.3

Plot the daily log returns in red line.

```{r}
plot(SPY.log.return, type = "l", col = "red", main = "Daily Log Returns for SPY")
```

# Question 2: (20 points)

## 2.1

Calculate the skewness and kurtosis of the SPY daily log return from Question 1, for both adjusted and unadjusted ones.

```{r}
library(moments)

# Adjusted -
SPY.Adjusted.log.return <- log(Cl(SPY))
SPY.Adjusted.log.return <- na.omit(SPY.Adjusted.log.return)
SPY.Adjusted.log.return.skew <- skewness(SPY.Adjusted.log.return)
SPY.Adjusted.log.return.kurtosis <- kurtosis(SPY.Adjusted.log.return)

# Unadjusted - 
SPY.log.return <- diff(log(Cl(SPY)))
SPY.log.return <- na.omit(SPY.log.return)
SPY.log.return.skew <- skewness(SPY.log.return)
SPY.log.return.kurtosis <- kurtosis(SPY.log.return)

```

## 2.2

Report the results in 2.1 using a 2×2 table (either data frame or matrix) such that: The column names are "SPY.skewness" and "SPY.kurtosis". And the row names are "Unadjusted" and "Adjusted".

```{r}
results_table <- matrix(c(SPY.log.return.skew, SPY.log.return.kurtosis,
                          SPY.Adjusted.log.return.skew, SPY.Adjusted.log.return.kurtosis), 
                        nrow = 2, ncol = 2, byrow = TRUE)

colnames(results_table) <- c("SPY.skewness", "SPY.kurtosis")
rownames(results_table) <- c("Unadjusted", "Adjusted")

results_table
```

# Question 3: (50 points)

## 3.1

Download options prices for ticker "SPY" for all expiration dates.

```{r}
SPY.options <- getOptionChain('SPY',NULL)
# head(SPY.options)
```

## 3.2

For calls and puts of each expiration date, add a column of "Price", which is the average of "Bid" and "Ask".

```{r}
for (i in 1:length(SPY.options)){
  SPY.options[[i]]$calls$Price <- mean(c(SPY.options[[i]]$calls$Bid, SPY.options[[i]]$calls$Ask))
  SPY.options[[i]]$puts$Price <- mean(c(SPY.options[[i]]$puts$Bid, SPY.options[[i]]$puts$Ask))
}

#head(SPY.options)
```

## 3.3

For calls and puts of each expiration date, add a column of "ImpliedVol", which is the implied volatility of the corresponding options calculated from root finding methods.

```{r}

getSymbols("SPY", from = "2020-01-01", to = Sys.Date())
SPY.S0 <- tail(SPY$SPY.Adjusted, n = 1)
r <- 0.07 * 0.01
SPY.options <- getOptionChain("SPY")
# strike price
(SPY.K1 <- SPY.options$calls$Strike[1])

# option price
(SPY.Price1 <- 0.5*(SPY.options$calls$Bid[1] + SPY.options$calls$Ask[1]))

# time to maturity
(T1 <- as.numeric(as.Date("2022-07-27") - Sys.Date())/252)


bs.call <- function(S0, K, T1, sigma, r){
  if(T1 <= 0){
    return(NA)
  }
  d1 <- (log(S0/K) + (r+0.5*sigma^2)*T1)/(sigma*sqrt(T1))
  d2 <- d1 - sigma*sqrt(T1)
  S0*pnorm(d1) - exp(-r*T1)*K*pnorm(d2)
}


f <- function(x){
  x^2 - 2
}

bisection.new <- function(f, a, b, tol = 0.001, N.max = 100){
  f.a <- f(a)
  f.b <- f(b)
  if(is.na(f.a*f.b) || f.a*f.b > 0){
    return(NA)
  }else if(f.a == 0){
    return(a)
  }else if(f.b == 0){
    return(b)
  }else if(is.na(f.a*f.b)){
    return(NA)
  }
  for(n in 1:N.max){
    c <- (a + b)/2
    f.c <- f(c)
    if(f.c == 0 || abs(b - a) < tol){
      break
    }
    if(f.a*f.c < 0){
      b <- c
      f.b <- f.c
    }else{
      a <- c
      f.a <- f.c
    }
  }
  return(c)
}

# implied.vol.call(S0,K,T1,r,10)
implied.vol.call <- function(S0, K, T1, r, price, method="bisection"){
  price.diff <- function(sigma) {
    bs_price <- bs.call(S0, K, T1, sigma, r)
    if (is.na(bs_price)) {
      return(1e20)
    } else {
      return(bs_price - price)
    }
  }
  if(method == "bisection"){
    return(bisection.new(price.diff, 0.01, 5))
  }
}
# implied.vol.call(SPY.S0,SPY.K1,T1,r,SPY.Price1)

# Add a column of implied volatility for calls and puts of each expiration date
# SPY.options <- getOptionChain("SPY", NULL)# all options
SPY.expiration <- names(SPY.options)# all expiration dates
T.vec <- (as.Date(SPY.expiration,"%b.%d.%Y")-Sys.Date())/365# calendar day
T.vec <- as.numeric(T.vec)# all time to maturities

for(i in 1:length(SPY.options)){
  SPY.options[[i]]$calls$Price <- 0.5*(SPY.options[[i]]$calls$Bid
                                           + SPY.options[[i]]$calls$Ask)
  for(j in 1:nrow(SPY.options[[i]]$calls)){
    SPY.options[[i]]$calls$impliedVol[j] <-
      implied.vol.call(SPY.S0,
                       SPY.options[[i]]$calls$Strike[j],
                       T.vec[i],
                       r,
                       SPY.options[[i]]$calls$Price[j])
  }
  SPY.options[[i]]$calls <-
    SPY.options[[i]]$calls[c("Bid", "Ask", "Strike","Price","impliedVol")]
}

for(i in 1:length(SPY.options)){
  SPY.options[[i]]$puts$Price <- 0.5*(SPY.options[[i]]$puts$Bid
                                           + SPY.options[[i]]$puts$Ask)
  for(j in 1:nrow(SPY.options[[i]]$puts)){
    SPY.options[[i]]$puts$impliedVol[j] <-
      implied.vol.call(SPY.S0,
                       SPY.options[[i]]$puts$Strike[j],
                       T.vec[i],
                       r,
                       SPY.options[[i]]$puts$Price[j])
  }
  SPY.options[[i]]$puts <-
    SPY.options[[i]]$puts[c("Bid", "Ask", "Strike","Price","impliedVol")]
}


#head(SPY.options)

```

## 3.4

Choose 3 expiration date for put options, plot volatility smiles

```{r}
plot(NA, xlim = c(300,500), ylim = c(0,0.4), xlab = "Strike",
     ylab = "ImpliedVol") 
lines(SPY.options[[14]]$calls$Strike,
      SPY.options[[14]]$calls$impliedVol,col = "red")
lines(SPY.options[[19]]$calls$Strike,
      SPY.options[[19]]$calls$impliedVol,col = "green")
lines(SPY.options[[21]]$calls$Strike,
      SPY.options[[21]]$calls$impliedVol,col = "blue")
legend("bottomleft", SPY.expiration[c(14,19,21)],
       fill = c("red","green","blue"))
```

## 3.5

Keep fields "Strike","Bid","Ask", "Price", and "ImpliedVol" and save the calls and puts of each expiration date in .csv file. Submit one of the .csv file also.

```{r}

for (i in 1:length(SPY.options)){
	SPY.options[[i]]$calls <- SPY.options[[i]]$calls[c('Strike', 'Bid', 'Ask', 'Price', 'impliedVol' )]
	SPY.options[[i]]$puts <- SPY.options[[i]]$puts[c('Strike', 'Bid', 'Ask', 'Price', 'impliedVol' )]
}

Exp <- names(SPY.options)
today <- format(Sys.Date())

setwd('C:\\Users\\amand\\OneDrive\\Desktop\\FE 515 (R)\\Final')

for(i in 1:length(Exp)){
	write.csv(SPY.options[[i]]$calls,file=paste0("SPYdata",today,"Exp",Exp[i],"calls.csv",sep=""))
	write.csv(SPY.options[[i]]$puts,file=paste0("SPYdata",today,"Exp",Exp[i],"puts.csv",sep=""))
	}
```
